---
title: "Gov 50 Final Project"
author: Keren Elmore
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---

## Milestone 3: 1st Data Visualization
```{r, echo = FALSE}
library(tidyverse)
library(ggplot2)
library(dplyr)
```

```{r, include = FALSE}
library(readr)
political_violence_data <- read_csv("~/Downloads/dataverse_files (1)/PoliticalViolence_May21 Client.csv")
```
```{r, include = FALSE}
political_violence <- political_violence_data |>
    select(-c(starttime, endtime, parentyn_1, parentyn_2, parentyn_3, parentyn_98))
political_violence
```

```{r, include = FALSE}
political_violence <- political_violence |>
  rename(moral_just = ucl_1_1, moral_strategy = ucl_1_2, terrorism = ucl_1_3)
```
```{r, include = FALSE}
print(political_violence)
```
```{r, echo = FALSE}
no_inj_morality <- political_violence |>
  filter(ideo5 == 1:5, experiment_split == c(1, 3, 9, 11)) |>
  mutate(
    ideology = case_when(
      ideo5 == 1 | ideo5 == 2 ~ "Liberal",
      ideo5 == 3 ~ "Moderate",
      ideo5 == 4 | ideo5 == 5 ~ "Conservative"
  )) |>
  mutate(
    experiment_ideo_split = case_when(
      experiment_split == 1 ~ "Pro abortion",
      experiment_split == 3 ~ "Anti abortion",
      experiment_split == 9 ~ "Pro immigration",
      experiment_split == 11 ~ "Anti immigration")
  ) |>
  ggplot(mapping = aes(x = experiment_ideo_split, y = moral_just, fill = ideology)) +
  geom_boxplot() +
  scale_fill_manual(values = c(Liberal = "steelblue1", Conservative = "indianred1", Moderate = "darkgoldenrod1")) +
  labs(
    x = "Political Violence Scenario Type (No Injuries)",
    y = "Morally Justifiable?")

no_inj_morality
```
The following data visualization between "experiment_split" and "moral_just" (relative to "ideo5") evaluates how conservative, liberal, and moderate leaning respondents view non-violence scenarios of political violence (regarding abortion and immigration) as morally justifiable. 

1 = Strongly Agree

2 = Agree

3 = Somewhat Agree

4 = Neutral

5 = Somewhat Disagree

6 = Disagree

7 = Strongly Disagree

-------

## Project thoughts

I am interested in exploring data related to housing in a major city (such as New York), specifically disparities among public housing residents, renters, and home owners. Exploring this data might lead me writing about how housing status (public housing vs renting vs ownership) intersects with race, income, or political affiliation. I would also be interested in thinking about how housing status for families specifically impacts education access and success in children. 

Some other thoughts include:

- Exploring correlations between landownership (specifically in Black or indigenous communities) and climate change

- Exploring companies that follow ESG (Environmental, social, and corporate governance) and the correlation to diversity optimization and environmental mitigation (climate change reduction). 

## Project Proposal
What is the relationship between political ideology and perspectives on political violence? In a time of heightened political tension and discourse, I want to explore how events of political violence (or the narratives formed around them) impact ideology and voting patterns. I hypothesize that those who have experience instances of political violence have generally more conservative views than those who have not. However, non-violent exposure contentious policy issues (such as abortion) fosters more liberal views. This is because typical while violence breeds thoughts of preventive action, non-violent political encounters breed feelings of advocacy and activism. A woman who chose to have an abortion will most likely have more pro-abortion views. My research will focus mostly on the "scenario_type" (measured on a scale from 1 to 12), "political party" (measured on a scale from 1 to 5), and "political ideology" (measured on a scale from 1 to 6) variables. These explanatory variables will be bridged with the outcome variable of "presvote20post". If a 'Pro abortion, no injuries' = 'Pro abortion' stance in "scenario_type" leads to voting for a Democrat presidential candidate, this would support my hypothesize while the adverse would not. 


```{r print-code, ref.label=knitr::all_labels(), echo = TRUE, eval=FALSE}

```

